\documentclass[boxes]{gsypset}

% Info for header
\mailbox{}
\initials{}
\collaborators{}
\class{Math 65}
\assignment{HW 9}
\duedate{May 27, 2016}

\usepackage{hyperref}

\newcommand{\pn}[1]{\left( #1 \right)}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\bk}[1]{\left[ #1 \right]}
\newcommand{\m}[1]{\begin{bmatrix} #1 \end{bmatrix}}
\newcommand{\f}[2]{\frac{#1}{#2}}
\renewcommand{\bf}[1]{\mathbf{#1}}
\renewcommand{\labelenumi}{{(\alph{enumi})}}

\begin{document}

\problemlist{1, 2, 3}

\begin{problem}
	Consider the inhomogeneous linear system of differential equations
	\[
		\bf{x}'=\m{5&-4\\6&-5}\bf{x}+\m{e^{t}\\0}.
	\]
	Note that the system matrix has eigenvalues $\lambda_1=1$ and
	$\lambda_2=-1$ with corresponding eigenvectors $\bf{v}_1=\m{1\\1}$
	and $\bf{v}_2=\m{2\\3}$.
	\begin{subproblems}
		\subproblem 
			Assume that the solution to this system $\bf{x}$ has the form
			$\m{a_1e^{-t}+b_1e^{t}+c_1te^{t}\\a_2e^{-t}+b_2e^{t}+c_2te^{t}}$
			for constants $a,b,c,d,e\mbox{ and }f$.  Use the method of
			undetermined coefficients to find the general solution to
			this system.
			\begin{solution}
				
			\end{solution}
		\subproblem Why was the assumption made in part (a) reasonable given
			what you know about the matrix $A$?
			\begin{solution}
				
			\end{solution}
		\subproblem Recalculate the solution using the integrating factor
			formula.  Reconcile your answer from this calculation with
			your previous calculation to show they are equivalent.
			\begin{solution}
				
			\end{solution}
	\end{subproblems}
\end{problem}

\begin{problem}
	Solve the initial value problem
	\[
		\bf{x}'(t)=\m{0&2\\-2&0}\bf{x}(t)
		+\m{\cos(\omega t)\\-\sin(\omega t)}
	\]
	where $\omega$ is a nonzero constant and
	$\bf{x}(0)=\m{a\\b}$. Make sure your solution is valid for all
	values of $\omega\neq0$.  In particular, be careful not to divide
	by 0 when $\omega=-2$.  Work out the solution in the case when
	$\omega\neq-2$, then work out the solution in the case when
	$\omega=-2$.  You may use either the undetermined coefficients or
	integrating factor method.  You will get extra credit if you
	perform both calculations and show that they are equivalent.\\\\
	\textbf{Hint:} If you're using the undetermined coefficients
	method, think about what we did in Math 45 when the forcing
	function took on the same form as the homogeneous solution.
\end{problem}
\begin{solution}
	
\end{solution}

\begin{problem}
	As we will discuss in class on Friday, closed-form solutions to
	non-constant-coefficient linear systems of differential equations
	are generally unavailable.  However, if you are lucky enough to
	solve a homogeneous (unforced) non-constant-coefficient linear
	system, solving the inhomogeneous (forced) version is relatively
	easy.  In this exercise, we'll walk you through how this works.\\\\
	Consider the following initial value problem:
	\[
		\bf{x}'(t)=A(t)\bf{x}+\bf{F}(t)=\m{2&-2e^{-t}\\e^{t}&0}\bf{x}
		+\m{4\\6}\qquad\text{with}\qquad\bf{x}(0)=\m{-1\\1}.
	\]
	\textbf{Hint:} Throughout this problem, you should use the formula
	\[
		\m{a&b\\c&d}^{-1}=\f{1}{ad-bc}\m{d&-b\\-c&a}
	\]
	to calculate the inverse of a $2\times2$ matrix. (Unfortunately,
	there are no simple formulas for larger matrices.)
	\textbf{Note:} To learn more about fundamental matrices and the
	derivation of the variation of parameters formula, read Section
	7.9 of Boyce and diPrima.
	\begin{subproblems}
		\subproblem First, verify that
			\[
				\bf{x}_h(t)=c_1\m{2e^{t}\\e^{2t}}+c_2\m{1\\e^{t}}
			\]
			is the general solution to the homogeneous version of the ODE
			(that is, the version of the ODE without the forcing term
			$\bf{F}$) for any $c_1$ and $c_2$.
			\begin{solution}
				
			\end{solution}
		\subproblem Arrange the two linearly independent solutions from part
			(a) as columns in a matrix $\Psi(t)$, which is called a
			\textit{fundamental matrix}. Verify that
			$\Psi'(t)=A(t)\Psi(t)$.
			\begin{solution}
				
			\end{solution}
		\subproblem The general solution to homogeneous equation is
			$\bf{x}_h(t)=\Psi(t)\bf{c}$ for some constant vector
			$\bf{c}=\m{c_1\\c_2}$. What must $\bf{c}$ be to satisfy the
			initial condition $\bf{x}_h(t_0)=\bf{x}_0$?
			\begin{solution}
				
			\end{solution}
		\subproblem The \textit{transition matrix} for this system is
			defined to be $\Phi(t,s)=\Psi(t)\Psi(s)^{-1}$. Calculate it.
			\textbf{Note:} The reason why we call this a ``transition
			matrix'' is that the effect of multiplication by the
			matrix $\Phi(t,t_0)$ is to transition the homogeneous
			solution forward in time from $t_0$ to $t$.
			\begin{solution}
				
			\end{solution}
		\subproblem Armed with the transition matrix corresponding to the
			homogeneous ODE, we can now calculate the solution to the
			inhomogeneous ODE using the variation of parameters formula
			\[
				\bf{x}(t)=\Phi(t,0)\bf{x}(0)
				+\int_{0}^{t}\Phi(t,s)\bf{F}(s)~ds.
			\]
			Use this formula to calculate the solution to the IVP.
			\begin{solution}
				
			\end{solution}
	\end{subproblems}
\end{problem}
\end{document}
